---
title: "RWorksheet_Redosendo#3b"
author: "Kyle Edward B. Redosendo"
date: "2025-11-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# 1. Create a data frame using the table below

## a. Write the codes
```{r}
df <- data.frame(
  Respondents = 1:20,
  Sex = c(2,2,1,2,2,2,2,2,2,2,1,2,2,2,2,2,2,2,1,2),
  Fathers_Occupation = c(1,3,3,3,1,2,3,1,1,1,3,2,1,3,3,1,3,1,2,1),
  Persons_at_Home = c(5,7,3,8,5,9,6,7,8,4,7,5,4,7,8,8,3,11,7,6),
  Siblings_at_School = c(6,4,4,1,2,1,5,3,1,2,3,2,5,5,2,1,2,5,3,2),
  Types_of_Houses = c(1,2,3,1,1,3,3,1,2,3,2,3,2,2,3,3,2,3,3,2)
)
```

## b. Describe the data
```{r}
print("Data Frame:")
df

print("Structure of the Data:")
str(df)

print("Summary of the Data:")
summary(df)
```

## c. Is the mean number of siblings attending 5?
```{r}
sibmean <- mean(df$Siblings_at_School)
print(paste("Actual mean is:", sibmean))
sibmean == 5
```

## d. Extract the 1st two rows
```{r}
subset_d <- df[1:2, ]
subset_d
```

## e. Extract 3rd and 5th rows with 2nd and 4th columns
```{r}
subset_e <- df[c(3, 5), c(2, 4)]
subset_e
```

## f. Select Types_of_Houses into a vector
```{r}
types_houses <- df$Types_of_Houses
types_houses
```

## g. Select male respondents whose father is a farmer
```{r}
subset_g <- df[df$Sex == 1 & df$Fathers_Occupation == 1, ]
subset_g
```

## h. Select female respondents with ≥ 5 siblings attending school
```{r}
subset_h <- df[df$Sex == 2 & df$Siblings_at_School >= 5, ]
subset_h
```

# 2. Create an empty data frame
```{r}
df_empty <- data.frame(
  Ints = integer(),
  Doubles = double(),
  Characters = character(),
  Logicals = logical(),
  Factors = factor(),
  stringsAsFactors = FALSE
)

print("Structure of the empty dataframe:")
str(df_empty)

write.csv(df_empty, "HouseholdData.csv", row.names = FALSE)
```

## a. Describe the results
The empty dataframe contains the correct column names and data types but **zero rows**.

# 3. Import the CSV file
```{r}
house <- read.csv("HouseholdData.csv", header = TRUE)
print("Imported CSV Data:")
house
```

## b. Convert Sex (Male=1, Female=2)
```{r}
house$Sex <- factor(house$Sex, levels = c("Male", "Female"), labels = c(1,2))
house$Sex
```

## c. Convert Types_of_Houses (Wood=1, Concrete=2, Semi-Congrete=3)
```{r}
house$Types_of_Houses <- factor(house$Types_of_Houses, levels = c("Wood", "Concrete", "Semi-Congrete"), labels = c(1,2,3))
house$Types_of_Houses
```

## d. Convert Fathers_Occupation (Farmer=1, Driver=2, Others=3)
```{r}
house$Fathers_Occupation <- factor(house$Fathers_Occupation, levels = c("Farmer", "Driver", "Others"), labels = c(1,2,3))
house$Fathers_Occupation
```

## e. Select all females whose father is a driver
```{r}
subset_3e <- subset(house, Sex == 2 & Fathers_Occupation == 2)
subset_3e
```

## f. Select respondents with ≥ 5 siblings attending school
```{r}
subset_3f <- subset(house, Siblings_at_School >= 5)
subset_3f